---
title: UQRUG 27
date: '2022-06-01'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2022-06-01: UQRUG 27

#### Attendees

* **Stéphane**: Library, here to help!
* **Emma**: looking for some help editing a graphic with ggplot2
* **Rene**: just tagging along
* **Laura**: working on GLMs
* **Leonie**: both tagging along and hoping for some help with the adehabitat package
* **Chris**: just tagging along
* **David**: using R on HPC 
* **Astrid**: R Markdown issues
* **Olalekan**: just tagging along
* 
* ...and 6 other UQRUGers!


#### Topics discussed and code

* ggplot2 customisation: moving legend, filtering data out
    * [Cédric Scherer's slides](https://www.cedricscherer.com/slides/OutlierConf2021_ggplot-wizardry.pdf) (with customisation of legend using the `guides()` function)


```{r}
library(dplyr)
# do not keep these three eye colours
starwars %>% 
  filter(!eye_color %in% c("blue", "yellow", "green"))
```

* Preparation for species distribution modelling. Convert dataframe to sf object with `st_as_sf()`, and will probably need to go from vector data to raster data with `terra::rasterize()`
    * The [CRAN Task View on spatial data](https://CRAN.R-project.org/view=Spatial) lists a lot of useful packages
* Importing spatial points for dolphin occurences, using sf. Constructing a convex hull from them and visualising on an interactive map:


```{r}
# read CSV as dataframe
dolph <- read.csv("Adehabitat.csv")

library(sf)
# convert the dataframe to an sf object
dolph_sf <- st_as_sf(dolph, coords = c("Longitude", "Latitude"))
# see it with default plot method
plot(dolph_sf)

# interactive map
library(tmap)
tmap_mode("view")
tm_shape(dolph_sf) +
  tm_dots()

# convex hull
dolph_hull <- st_convex_hull(st_union(dolph_sf))

# visualise both
tm_shape(dolph_hull) +
  tm_borders() +
tm_shape(dolph_sf) +
  tm_dots()
```

* Detecting anomalies in chronological sequence of a dataframe. `dplyr::lag()` and `dplyr::lead()` functions can be used for comparisons. `any()` and `all()` help reducing many logical values to one.
* R Markdown troubles: Rmd is self-contained and needs to include all the necessary code. Its working directory is by default the directory where the .Rmd file is saved.
* factoextra's `fviz_pca*()` functions for PCA, colouring points per group.
    * [STHDA has examples](http://www.sthda.com/english/wiki/fviz-pca-quick-principal-component-analysis-data-visualization-r-software-and-data-mining)


#### Links

* [R Ladies Brisbane YouTube channel](https://www.youtube.com/channel/UC9oOCJe8kwkZ_6IgTmet9oQ)
* [Geospatial Analysis Community of Practice](https://geospatial-community.netlify.app/)


