---
title: UQRUG 10
date: '2020-10-19'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2020-10-19: UQRUG 10

#### Attendees and problems

Please add your name and the problem you'd like some help with!

* **Stéphane** (Library): here to help! 
* **Chantelle**: Biological Sciences. I can't stay to long as I have to go track carpet pythons at 4pm ;)
* **Einat** (Social Sciences, R studio beginner): Apologies in advance would have to leave early. have some issues wiTh minor and basic procedures, like ...CONDITIONS!! and also with reading labels from csv/excel file. 
* **Patrick** (Agricultural Science) Honours student
* **Gabriel**: Hydrologist, spatial data analyst. PhD candidate at SMI. R beginner and Python enthusiast !
* **Natalie** (Mathematics): Undergraduate Mathematics and Statistics student
* **Phoebe**: learn more about data viz!
* ... and few more!


#### Today's challenge

Try playing with the latest TidyTuesday dataset: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-10-13/readme.md

Share your findings and cool visualisations with the group!


#### Recoding data

This is a working example of recoding using dplyr's `case_when()`:


```{r}
library(dplyr)
CO2 %>% 
  mutate(uptake = case_when(
    uptake > 12 ~ "yay!",
    uptake > 10 ~ "OK.",
    TRUE ~ "oh no..."
  ))
```

The `TRUE` line is for "everything else".


#### Finding files

The "working directory" is the default location:


```{r}
getwd() # know where you are
setwd("path/to/correct/location") # change it

# functions will look there by default
read.csv("filename.csv")
# but you can use longer, absolute filepaths
read.csv("C:/Users/myname/filename.csv")
```

If you don't want to deal with that: **use R Projects**! Creating an R Project will set the working directory for you.


#### Dealing with overflowing labels

```{r}
library(ggplot2)
ggplot(diamonds, aes(x = cut)) +
  geom_bar()

# horizontal bar chart
ggplot(diamonds, aes(y = cut)) +
  geom_bar()

# abbreviate the labels
abbreviate(diamonds$cut)

# using it in the visualisation is nice and concise
ggplot(diamonds, aes(x = cut)) +
  geom_bar() +
  scale_x_discrete(labels = abbreviate)
```

On the other hand, if labels are overflowing inside the canvas, we can expand axis limits:


```{r}
# expand limits
ggplot(diamonds, aes(x = cut)) +
  geom_bar() +
  ylim(c(0, 25000))
```

However, ggplot2 should expand the plot area to include all the geometries automatically, including a `geom_label()` or `geom_text()`.


#### Shared resources

Resources shared during the meetup can be listed here:

* Packages:
    * epiDisplay: https://rdrr.io/cran/epiDisplay/
    * lme4: https://www.rdocumentation.org/packages/lme4/versions/1.1-23/topics/lme4-package
    * Seurat: https://cran.r-project.org/web/packages/Seurat/index.html
    * reshape2: https://cran.r-project.org/web/packages/reshape2/index.html
* Functions:
    * `case_when()` to recode a variable: https://dplyr.tidyverse.org/reference/case_when.html
* Data viz:
    * ggplot2 extensions: https://exts.ggplot2.tidyverse.org/gallery/
    * Data to Viz: https://www.data-to-viz.com/
    * lattice: https://www.rdocumentation.org/packages/lattice/versions/0.20-41



