---
title: UQRUG 14
date: '2021-03-15'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2021-03-15: UQRUG 14

#### Participants

* **Stéphane Guillou** (Library): ...
* **Paula Andrea Martinez** (ReSA): are you interested in a Markdown workshop? two hours on the 24th of March by [NeSI](https://www.eventbrite.co.nz/e/rmarkdown-for-researchers-weave-together-narrative-text-and-code-registration-144069029345)
* **Fathin Azizan** (HDR SAFS): Need help to check on coding for the multiple linear regression using raster data.
* **Robyn**
* **Phoebe**: Hi everyone!
* **Tmnit**
* **Violeta**
* **Patrick**
* **Aljay**


#### This week's TidyTuesday dataset

[Bechdel Test](https://en.wikipedia.org/wiki/Bechdel_test) dataset: https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-09/readme.md

Get the data and play with it:


```{r}
raw_bechdel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/raw_bechdel.csv')
movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-09/movies.csv')
```

Share code and pictures here!


#### Shared resources

* Intro slides by Grant McDermott: https://raw.githack.com/uo-ec607/lectures/master/04-rlang/04-rlang.html#1
* mapsf is the successor to cartography: https://rgeomatic.hypotheses.org/2212
* Open Science Foundation: https://osf.io/
* OpenStreetMap data:
    * Geofabrik to get bulk OpenStreetMap downloads: http://download.geofabrik.de/
    * osmextract package: https://cran.r-project.org/web/packages/osmextract/index.html
    * osmdata package: https://docs.ropensci.org/osmdata/



