---
title: UQRUG 25
date: '2022-03-30'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2022-03-30: UQRUG 25

#### Attendees

* **Steph** (Library): helping out!
* **Vicki Martin**: Postdoc, SEES
* **Nisa Abeysinghe**
* **Richard Bell**: PhD, POLSIS
* **Chris Mancini**: HDR - MPhil, School of Civil Engineering
* ... and 6 more UQRUGers!


#### Topics discussed and code

##### Big raster files

Use terra instead of raster, and use a temporary directory:


```{r}
library(terra)
terraOptions(tempdir = "path to somewhere with lots of space")
as.points() # instead of rasterToPoints()
```

The [Research Computing Centre](https://rcc.uq.edu.au/) can also provide access to supercomputers


##### [Patchwork](https://patchwork.data-imaginist.com/) for joining plots made with [ggforestplot](https://nightingalehealth.github.io/ggforestplot/articles/ggforestplot.html)

To merge two plots with the same y-axis. After creating Forest plot 1 (with y-axis labels), create Forest plot 2 with y-axis labels removed:


```{r}
library(ggforestplot)
F2plot <- 
  forestplot(
    df = F2,
    name = term,
    estimate = estimate,
    se = std.error,
    pvalue = p.value,
    psignif = 0.05, 
    title = "Plot 2",
    xlab = "estimate",
    ylab = element_text(family = "", size = 10)
  )+ 
  theme(axis.text.y = element_blank())
```

Or try replacing text in ylab (above) to ylab = ""

Then merge the plots:


```{r}
library(patchwork)
F1plot / F2plot
```

##### Slow `check_model()`

`check_model()` in [performance](https://easystats.github.io/performance/) package: weird behaviour in R Markdown, takes too long to execute. Richard might come back later with the answer!


