---
title: UQRUG 40
date: '2023-07-26'
description: '**Overview**: Using quarto and revealjs to create presentations with R markdown. <br /> **Questions**: learning about knitr, combining shiny with quarto, fixing OneDrive breaking R, cmdstanr, brms'
categories: meeting
toc: true
toc-depth: 5
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

## 2023-07-26: UQRUG 40

### R Overview of the Month

Nick  will be providing an overview on creating **HTML based presentations in R with [Quarto](https://quarto.org/docs/presentations/) and [reveal.js](https://revealjs.com/)**

Find more details and the slides here:
<https://uqrug.netlify.app/posts/2023-07-26-july-q-pres/>

### Attendees
Add your name, where you're from, and why you're here:

| Name | Where are you from? | What brings you here? |
|--------|--------|--------|
| Nicholas Wiggins | Library | Here to help! |
| Mikesh Patel | QCMHR | Learning R |
| Raul Riesco | ACE | Learning/helping |
| Ryan | SOE | Learning |
| Luke Gaiter | Library | Here to help! |
| Rene Erhardt | PACE |learning||
| Trinh Huynh |USC|learning|
| Danelle Kenny|CHSR|checking in and trying new things|
| Pia Wohland| SENV|Checking in|
| Jaimee Wheelhouse ||Very new to R|
| Graeme Mattison | TPCH | Learning R!|
| Fran Powell| SVS| Need some help |
| Jocelyn Hume | UQCCR | Still learning|
| David Green|UQ RCC|learning/helping ... R on HPC and Shiny|
|Anne-Claire Bouton | SOE | Learning |
| Windy | SENV | Still learning
| Valentina Urrutia Guada | Library | Here to help!
| Gerry Katai ||    new to R learning 


### Questions

#### Q1 - Learning more about knitr - Trinh Huynh
I haven't worked with R markdown and I would like to learn from beginning- Trinh Huynh

**Answers**

*   Take a look at this episode of the R for Reproducible Science workshop:
    * <https://swcarpentry.github.io/r-novice-gapminder/15-knitr-markdown.html>
* The library offers introductory training modules, as well as advanced. You can look for the upcomming sessions in this website:
  * <https://web.library.uq.edu.au/library-services/training>
* After doing that training, you're more than welcome to bring questions that you have to UQRUG

---

#### Q2 - knit Shiny/plotly objects to local html - Mikesh
Solutions for saving Rmarkdown files with shiny objects locally (no server)? 
Want to allow someone without R installed on their computer, access the interactive html file. maybe todays session?...

**Answers**

* Nick - I haven't tried combining shiny and quarto, as Shiny will always need that Shiny server running, so I try to stick to HTML related dependencies.
* If you want to have a quarto website or presentation, you can easily publish them to a variety of free sites such as GitHub Pages, Netlify, or Quarto Pub:
    * <https://quarto.org/docs/publishing/>
*  @Mikesh When you are thinking about this data access question again, please reach out to me via rcc-support@uq.edu.au

---

#### Q3 - How to get cmdstanr working - Fran
When I run cmdstan R, I get a make error.
```{r}
cmdstanr:::get_cmdstan_flags('STANCFLAGS')
Error in `(function (command = NULL, args = character(), error_on_status = TRUE, …`:
! System command 'make' failed
---
Exit status: 2
Stderr:
make: *** /c/Users/uqfpowe1/OneDrive: Is a directory.  Stop.
---
```

**Answers**

*  We spent a while working on this one. Eventually David, Valentina, Nick, and Fran managed to get it working.
*  The issue appears to be that cmdstanr is for some reason pointing at the OneDrive, rather than the local C: drive
*  We tried changing the standard working directory, but this did not work
*  We also pulled everything off of OneDrive that was associated with R and cmdstanr, but it still wasn't working
*  David got us to open the makefile command file inside the cmdstanr installation folder and we found that it was hard coded to run from the $HOME directory.
*  This is apparently different to the working directory. We found that we could check what this was with the following code:
    * `Sys.getenv("HOME")`
* We were able to change this with the following code:
  * `Sys.setenv(HOME="C:/Users/MyName/Documents")`
* We then had the code working.
---
