---
title: UQRUG 15
date: '2021-04-19'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2021-04-19: UQRUG 15

#### Participants

* **Stéphane Guillou** (Library): I'd like start working on a package to query the MusicBrainz API
* **Catherine Kim** (Library/Biology): Write a function to automate coral reef image download from US NOAA database CoRIS.  Cleaning up code for sharing.
* **Rene Erhardt** (Pharmacy): beginner, just listening...
* **Andy Cui** (Engineering): First time attendance with aim of reshaping spectral data for visualisation purposes
* **Rhiannon Jeans** (Neuroscience): First time attending the meeting but not new to R. Looking to help others or learn from others. 
* ...

#### TidyTuesday datasets

See if you find an interesting dataset to play with: https://github.com/rfordatascience/tidytuesday#datasets

Share code and pictures here!


#### Shared resources

* R Markdown news: https://blog.rstudio.com/2021/04/15/2021-spring-rmd-news/
* reshape2 tutorial: https://www.datacamp.com/community/tutorials/long-wide-data-R
* Tidyverse information: https://www.tidyverse.org/
* ggplot2 visualisation wizardry: https://www.cedricscherer.com/slides/useR2021.pdf
* pipe operator:
    * first introduced by magrittr: https://github.com/tidyverse/magrittr
    * but coming to R base: https://cran.r-project.org/doc/manuals/r-devel/NEWS.html
* R Markdown gallery: https://rmarkdown.rstudio.com/gallery.html
* Tables
    * flextable: https://davidgohel.github.io/flextable/
    * officer: https://davidgohel.github.io/officer/
    * frequency (for tables similar to SPSS' `FREQUENCIES`): https://github.com/wilcoxa/frequency

For example, try this:


```{r}
library(dplyr)
library(frequency)
options(frequency_open_output = TRUE)
mtcars %>% 
  select(carb, cyl) %>% 
  freq()
```

