---
title: UQRUG 19
date: '2021-08-16'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2021-08-16: UQRUG 19

#### Participants

* **Stéphane Guillou** (Library): just here to help!
* **Catherine Kim** (Library/Biology): working on publishing code with a paper
* **Jordan Pennells** (ANOVA structure)
* **Violeta Berdejo-Espinola** (make my code more efficient)
* **Omkar Ravindranatha Katagi** (Here to learn R)
* ... and 8 more UQRUGers!


#### Shared resources

Links we talked about today:

* Great books / resources to get started:
    * ***R for Data Science*** (aka "R4DS"): https://r4ds.had.co.nz/
        * [Slack community of R4DS](https://rfordatascience.slack.com/join/shared_invite/zt-n46lijeb-2RRzQ70U34eH530~PyZsmg#/shared-invite/email)
        * On Twitter: https://twitter.com/R4DScommunity
    * **R Cookbook**: https://www.cookbook-r.com/
    * **RStudio Education**: https://education.rstudio.com and https://education.rstudio.com/learn/
    * **learnr** package (also integrated in the RStudio "tutorial" tab): https://rstudio.github.io/learnr/
    * **article on linked points between boxplots**: https://datavizpyr.com/how-to-connect-data-points-on-boxplot-with-lines/


#### Code snippets

Change the order of categorical variable levels (so ggplot2 uses that order instead of the alphabetical order):


```{r}
library(dplyr)
library(ggplot2)
                          
# relevel factors using forcats package
library(forcats)
f <-  factor(c("a", "b", "c", "d"), levels = c("b", "c", "d", "a"))
fct_relevel(f, "a", "c", "d", "b")

# relevel eye_color of starwars data

starwars %>% 
   mutate(eye_color = fct_relevel(eye_color, "yellow")) %>% 
   ggplot(aes(x = eye_color)) +
   geom_bar()
```

Use case_when to rate or recode data:


```{r}
library(dplyr)

starwars
range(starwars$height, na.rm = TRUE)

# just based on height
starwars_rated <-  starwars %>% 
   mutate(height_rating = case_when(height < 70 ~ "very small",
                                    height < 140 ~ "quite small",
                                    height < 200 ~ "medium",
                                    TRUE ~ "tall"))                                       
                                    
# recode based on multiple conditions
range(starwars$mass, na.rm = TRUE) # check range of mass, second vairable to recode by

starwars_rated <-  starwars %>% 
   mutate(height_rating = case_when((height < 70) & (mass < 700)~ "very small and light",
                                    height < 140 ~ "quite small",
                                    height < 200 ~ "medium",
                                    TRUE ~ "tall"))
          
```

