---
title: UQRUG 28
date: '2022-06-29'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2022-06-29: UQRUG 28

#### Attendees

* **Stéphane**: Library | here to help and say goodbye
* **Chris**: Civil Engineering - Transport | just tagging along
* **Luke**: Library | here to help and say hello
* **Olalekan** Biological Sciences | here to say hello...


#### Topics discussed and code

##### Iterating instead of using repetitive code

The trick here is to:

1. Encapsulate the repetitive code into a function, exposing the things that are likely to change as arguments
2. Create a vector of values (or several)
3. Use a for loop, or an `apply()` function, or a `map()` function (from purrr) to map the function to each element


```{r}
# iterating

?mean
# custom function
custom_mean <- function (numbers, trim_ratio) {
  # process the data
  my_mean <- mean(x = numbers, trim = trim_ratio, na.rm = TRUE)
  # save the it as file
  saveRDS(my_mean, file = paste0(numbers[1], trim_ratio, ".rds"))
}

# use the function
custom_mean(c(1,5,NA), 0.2)

# iterate
list_to_iterate_on <- list(c(1,5,NA),
     c(1,4,6,8),
     c(1,2,7,9,NA))

trim_ratios <- c(0.2, 0.7, 0.1)

library(purrr)
# alternative to apply functions or for loops
map2(list_to_iterate_on, trim_ratios, custom_mean)

# pmap(): construct a dataframe of all combinations,
# with each column containing the argument values to use

# construct the dataframe of all combinations
animal <- c("magpie", "pelican", "ibis")
treatment <- c("dry food", "sludge", "grain")
# only three rows
experiment <- data.frame(animal, treatment)
# all combinations
library(tidyr)
all_combinations <- expand(experiment, animal, treatment)

```

##### List files

```{r}
# listing files
list.files() # all files in working directory
only_rds <- list.files(pattern = "rds")
only_rds <- list.files("analysis", "rds")
only_rds
# full path (from working directory)
only_rds <- list.files("analysis", "rds", full.names = TRUE)
only_rds
```

##### Remove file extension from path

```{r}
# remove file extension from path
library(stringr)
str_replace("filename.txt", ".txt", "")
```

##### Extract information from filenames

Using the tidyverse and pdftools for preparing PDF text before analysis with quanteda. pdftools was used for its specific ability to return the page numbers of the pdfs.


```{r}
library(pdftools)
library(tidyverse)

# create a list of the PDF file paths
myfiles <- list.files(path = "./pdfs", pattern = "*.pdf", all.files = FALSE,
                      full.names = TRUE, recursive = TRUE,
                      ignore.case = FALSE, include.dirs = TRUE, no.. = FALSE)

# Function to import each pdf file, and place the text in a dataframe
import_pdf <- function(k){
    # turn the pdf into a text list each page will become a row
  pdf.text <- pdftools::pdf_text(k)
    # flatten the list
  pdf.text<-unlist(pdf.text)
  pdfdf <- data_frame(pdf.text)
    # turn the list into a dataframe, extracting the year from the path and using the separate function to extract the state from the path
  data_frame(pdf.text) %>%
    mutate(year = str_extract(k, "[:digit:]{4}") %>% as.integer(), pagenumber = row.names(pdfdf), filename = k) %>% 
    separate(filename, c(NA,NA,"state",NA), sep = "/")
  
}
    # run the function on all pdf files
all_pdfs<- map_dfr(myfiles, import_pdf)
```

See the [quanteda tutorials](https://tutorials.quanteda.io/)


