---
title: UQRUG 22
date: '2021-11-15'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2021-11-15: UQRUG 22

#### Attendees

* **Catherine**: Biology/Library, publish data for a paper 
* **Grechel**: PhD candidate/ PACE temporal trend analysis
* **Siu**: Bachelor of Biomedical Science
* **Einat**: PhD candidate / Civil Engineering School 


#### Code snippets

Example of for loop vs map() function from purrr package.


```{r}
# for loop to calculate the median of every column in mtcars
output <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {
   output[[i]] <- median(mtcars[[i]])
}
output # see the output

# same as above using a map() function
map_dbl(mtcars, median)
```

#### Shared resources

* See Stéphane's [tidyverse lesson](https://gitlab.com/stragu/DSH/-/blob/master/R/tidyverse_next_steps/tidyverse_next_steps.md) for more on purrr
* [ResBazQLD 2021](https://resbaz.github.io/resbaz2021/brisbane/) running 24-26th of Nov at UQ. Regsitration is $25.


