---
title: UQRUG 17
date: '2021-06-21'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2021-06-21: UQRUG 17

#### Participants

* **Stéphane Guillou** (Library): just here to share help!
* **Catherine Kim** (Library/Biology): playing around spatial and open datasets from this [World's Oceans Day R-blogger post](https://www.r-bloggers.com/2021/06/celebrating-world-ocean-day-ropensci-style/)
* ... and 8 more UQRUGers!


#### Shared resources

* World Ocean Day post: https://www.r-bloggers.com/2021/06/celebrating-world-ocean-day-ropensci-style/
* performance package, `check_model()` function to check model assumptions: https://rdrr.io/cran/performance/man/check_model.html
* Create custom ggplot2 themes: https://rpubs.com/mclaire19/ggplot2-custom-themes
* Regular expression + stringr cheatsheet: http://edrub.in/CheatSheets/cheatSheetStringr.pdf
    * Read more about regexes: https://en.wikipedia.org/wiki/Regular_expression
* stargazer package for nice tables:
    * On CRAN: https://cran.r-project.org/web/packages/stargazer/index.html
    * Manual: https://cran.r-project.org/web/packages/stargazer/stargazer.pdf
    * warning: takes dataframes(!), not tibbles! (Might silently fail)
* The sf package now uses a spherical geometry by default: https://r-spatial.github.io/sf/articles/sf7.html


#### Tips for newcomers

* Projects for everything! Using R Projects allows you to keep separate projects... separate! And find files easily. Start organised to stay organised.
* *R 4 Data Science* book: https://r4ds.had.co.nz/



#### Code snippets

##### Export to CSV

```{r}
# export with write.csv()
write.csv(ToothGrowth,
          file = "exports/tooth_growth.csv",
          row.names = FALSE, # remove rowname column
          na = "") # empty cell instead of "NA"
```

##### Missing data handling in `dplyr::filter()`

```{r}
# insert missing data:
ToothGrowth[27,3] <- NA
library(dplyr)
# dplyr filtering:
filtered_dplyr <- filter(ToothGrowth, dose > 0.5)
    # dplyr filter gets rid of rows returning NA
# base indexing:
filtered_base <- ToothGrowth[ToothGrowth$dose > 0.5, ]
    # base subsetting with [] keeps rows returning NA
# check documentation:
?dplyr::filter
```

> "[In `dplyr::filter`], note that when a condition evaluates to NA the row will be dropped, unlike base subsetting with `[`."


##### New `alt` argument in ggplot2

Great news for accessibility in the new ggplot2 version 3.3.4: https://cloud.r-project.org/web/packages/ggplot2/news/news.html

For example:


```{r}
library(ggplot2)
ggplot(ToothGrowth, aes(x = supp, y = len)) +
  geom_boxplot() +
  labs(alt = "Boxplot of tooth growth rate with different supplements.\n
       The length is generally higher with the OJ supplement.")
```

This will allow tools that create documents that include ggplot2 visualisations to make use of the alternative text, for screen readers for example.

Know that R Markdown has the chunk option `fig.alt` for that purpose as well.


