---
title: UQRUG 24
date: '2022-02-23'
categories: meeting
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval= FALSE)
library(tidyverse)

```

### 2022-02-23: UQRUG 24

#### Attendees

* **Stéphane Guillou**: just helping out
* **Svetlina Vasileva**
* **Trinh Huynh** (USC)
* **Richard Bell**
* **Luke Gaiter** (ggplot2 lesson)
* ...and 5 other UQRUGers!


#### Code

Convert a table to a `gt` object and then save it to RTF (which can be opened by Word).


```{r}
library(gt)
gt(mtcars) %>% 
  gtsave("test.rtf")
```

Attempting to highlight/label outliers in linear regression plot

```{r}
### geom_text_repel
# only label players with QoG >  1 or < 0.2
# align text vertically with nudge_y and allow the labels to 
# move horizontally with direction = "x"
ggplot(linear_regression, aes(x= QOG, y = ten_political_conflict, label = district)) + 
        geom_point(color = dplyr::case_when(linear_regression$QOG > 1 ~ "#1b9e77", 
                                            linear_regression$QOG < 0.2 ~ "#d95f02",
                                            TRUE ~ "#7570b3"), 
                   size = 3, alpha = 0.8) +
        geom_text_repel(data          = subset(linear_regression, QOG > 1),
                        nudge_y       = 32 - subset(linear_regression, QOG > 1)$QOG,
                        size          = 4,
                        box.padding   = 1.5,
                        point.padding = 0.5,
                        force         = 100,
                        segment.size  = 0.2,
                        segment.color = "grey50",
                        direction     = "x") +
        geom_label_repel(data         = subset(linear_regression, QOG < 0.2),
                         nudge_y       = 16 - subset(linear_regression, QOG < 0.2)$QOG,
                         size          = 4,
                         box.padding   = 0.5,
                         point.padding = 0.5,
                         force         = 100,
                         segment.size  = 0.2,
                         segment.color = "grey50",
                         direction     = "x") +
        scale_x_continuous(expand = expansion(mult = c(0.2, .2))) +
        scale_y_continuous(expand = expansion(mult = c(0.1, .1))) +
        theme_classic(base_size = 16)
```

#### Shared resources and topics discussed

* Using Cloudstor's SWAN: [documentation](https://support.aarnet.edu.au/hc/en-us/articles/360000575395-What-is-CloudStor-SWAN-)
* Export tables to DOC or DOCX:
    * Svet tried to use [arsenal's write2word() function](https://mayoverse.github.io/arsenal/reference/write2specific.html), but didn't work... inside a R Markdown chunk! Had to run it outside, possibly because the function itself uses knitr...
    * Richard suggested using [stargazer](https://cran.r-project.org/web/packages/stargazer/index.html)
    * [gt](https://gt.rstudio.com) is a powerful package for customised tables, and can export to RTF, and its website has a useful [list of R packages for creating tables](https://gt.rstudio.com/#how-gt-fits-in-with-other-packages-that-generate-display-tables).
* [forcats](https://forcats.tidyverse.org/index.html) has been used twice during the session, once for changing the order of the levels to something arbitrary with `forcats::fct_relevel()`, and once for ordering them by value (with `dplyr::arrange()` followed by `forcats::fct_inorder()`)
* Sustainable transport planning with [stplanr](https://docs.ropensci.org/stplanr/)
* [lidR](https://cran.r-project.org/web/packages/lidR/news/news.html) moves away from sp/raster and uses sf/terra/stars instead
* [gghighlight](https://cran.r-project.org/web/packages/gghighlight/vignettes/gghighlight.html) was suggested to easily highlight (an label) cases on ggplot2 visualisations


